---
title: "Project"
author: "Andreas Breuß"
date: "27 4 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


##```{r}

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.12")


BiocManager::install("affy")

BiocManager::install("vsn")


```{r}

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/tra/")

a=read.csv(file="human.tsv",sep="\t")

tiss=a[,11]

ind=which(tiss=="thyroid")

TRA.symbol=a[,3]

thyroid.TRA1=TRA.symbol[ind]

head(a)

```


```{r}
 
library(affy)
library(vsn)
library(AnnotationDbi)
library(hgu133plus2hsenstcdf, lib.loc = "C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/Packages/")
library(hgu133plus2hsenstprobe, lib.loc ="C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/Packages/")
library(hexbin)
library(rstudioapi)
library(tidyverse)
```

# used for setwd 
```{r}
# used for setwd() 

projectPath <- dirname(rstudioapi::getSourceEditorContext()$path)
projectPath

```



# Read in .CEL files
```{r}
#Kontrolle
#ind = which (new%in%c ()) 
##dev.copy2eps(file="QC_.....eps")
##setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/plots/")
#data.norm=vsnrma(data)

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/rawdata/GSE149507 lung cancer")

data = ReadAffy()
data@cdfName<-"HGU133PLUS2_Hs_ENST"

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/sessions/rda")

lungcancer_matrix <- as.data.frame(exprs(data))

save.image(file="rawdata.lungcancer.pdf")
image(data, col = rainbow(100, start = 0, end = 0.75)[100:1])

```


# overview
```{r}

#microarrays
dim(lungcancer_matrix)[2]

# names of the samples
colnames(data)
substr(colnames(lungcancer_matrix), 1,10)

# gene count
dim(lungcancer_matrix)[1]

# probe set names (or Affymetrix IDs) - are changed when changing cdfName
featureNames(data)[1:10]

#expression table
head(lungcancer_matrix)

```


# cleanup NA and/or space
```{r}
 
apply(lungcancer_matrix, 2, function(x){sum(is.na(x))})
apply(lungcancer_matrix, 2, function(x){sum(x == 0)})
sum(sapply(rownames(lungcancer_matrix), function(x){sum(x == "")}))

```


# normalization
```{r}
data.norm <- vsnrma(data)

head(data.norm)

#save

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/sessions/rda")

save.image(file="normalized_data.rda")

```

#Plots
```{r}

#meanSdPlot
meanSdPlot(data.norm)

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/plots/")
dev.copy2eps(file="meanSdPlot_lungcancer_vsnrma_normalized.eps")



# Boxplot before normalization
boxplot(data, col=rainbow(150),cex.axis=0.5, main="Gene expression in human breast cancer")

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/plots/")
dev.copy2eps(file="boxplot_lungcancer_rawdata.eps")

#Boxplot normalized data
boxplot(exprs(data.norm), col=rainbow(150),cex.axis=0.5, main="Gene expression in human breast cancer")

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/plots/")
dev.copy2eps(file="boxplot_lungcancer_vsnrma_normalized.eps")


```



#RNA degeneration plot
```{r}

rnadeg.raw = AffyRNAdeg(data)

#shift & scale
plotAffyRNAdeg(rnadeg.raw, col=rainbow(24), transform = "shift.scale")
title(sub = "lungcancer rawdata")


setwd(paste(projectPath, "plots" ,sep = "/"))
dev.copy2pdf(file = "lungcancer_rnadeg_rawdata.pdf")


#shift
plotAffyRNAdeg(rnadeg.raw, col = rainbow(24), transform = "shift.only")
title(sub = "lungcancer rawdata")


setwd(paste(projectPath, "plots" ,sep = "/"))
dev.copy2pdf(file = "lungcancer_rawdata.pdf")

```



#Scatterplot
```{r}

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/plots/")

plot(exprs(data.norm)[,c(1,2)],pch=".")+
abline(0,1,col="red")

dev.copy2eps(file="file.name")
dev.off()

head(data)
head(data.norm)

```


# Extract expression values
```{r}
dataExprs = exprs(data.norm)
head(dataExprs)
dim(dataExprs)
```

# ensemble IDs
```{r}
dataExprs = dataExprs[grepl("ENST", rownames(dataExprs)), ]
head(dataExprs)
dim(dataExprs)
```

# remove x_at suffix
```{r}
rownames(dataExprs) <- unlist(lapply(strsplit(rownames(dataExprs), split = "\\."), "[", 1))
head(dataExprs)
```

setwd("C:/Users/metho/iCloudDrive/Studium/4.FS/Bioinfo/RStudio/tables/")

data.matrix=exprs(data.norm)
head(data.matrix)


a=read.csv("ensembl.103.txt",sep="\t")

head(a)
affy.ensembl=as.character(a[,5])

symbol.ensembl=as.character(a[,4])

names(symbol.ensembl)=affy.ensembl

head(symbol.ensembl)

#look into data matrix

head(data.matrix)

affyIDs=rownames(data.matrix)

head(affyIDs)

symbol=symbol.ensembl[affyIDs]

head(symbol)

dim(data.matrix)

length(symbol)

#re-apply rownames to gene symbols in the data.matrix

rownames(data.matrix)=as.character(symbol)

head(data.matrix)

##This is the matrix you can calculate with##


dataExprs = exprs(data.norm)

head(dataExprs)
dim(dataExprs)

thyroid.TRA1 = rownames(dataExprs)

row.ind=which(thyroid.TRA1%in%symbol)

#or (please, note that sometimes upper case and lower case and other problems disturb the search, for example levels, which you can eliminate with as.character and 
#toupper


row.ind=which(toupper(thyroid.TRA1) %in% as.character(symbol))

data.sub=data.matrix[row.ind,]


par(las=2)

boxplot(t(data.sub),col=rainbow(length(rownames(data.sub))),main="gene expression of thyroid-specific genes in thyroid cancer",cex.axis=0.8)

#sort alphabetically

order.vector=sort(colnames(t(data.sub)),index.return=T)$ix

boxplot(t(data.sub)[,order.vector],col=rainbow(length(rownames(data.sub))),main="gene expression of thyroid-specific genes in thyroid cancer",cex.axis=0.8)

```
#Analysis

###```







